Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude

### 2016-11-08 RMS/final-version/ALL_15-TIE-3480.tex
final version: modified photos and renamed

%\documentclass[onecolumn]{IEEEtranTIE}
\documentclass[journal]{IEEEtranTIE}
\usepackage{graphicx}
\usepackage[noadjust]{cite}
\usepackage{picinpar}
\usepackage[cmex10]{amsmath}
\interdisplaylinepenalty=2500
\usepackage{stfloats}
\usepackage{url}
\usepackage{flushend}
\usepackage[latin1]{inputenc}
\usepackage{colortbl}
\usepackage{soul}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{color}
\usepackage{alltt}
\usepackage[hidelinks]{hyperref}
\usepackage{enumerate}
\usepackage{siunitx}
\usepackage{breakurl}
\usepackage{epstopdf}
\usepackage{pbox}

%% My packages
%% \usepackage[cmex10]{amsmath}
%% \interdisplaylinepenalty=2500
%% \usepackage{algpseudocode}
%% \usepackage{array}
%%\usepackage{dblfloatfix}
%%\usepackage[caption=false,font=footnotesize]{subfig}

%% \usepackage{tikz}
%% \usetikzlibrary{arrows.meta}
%% \usepackage{mathabx}

\newcommand{\hide}[1]{\ignorespaces}
\newcommand{\jx}[1]{{\bf Jiaxiang: }#1{ \bf End}}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}

% Define the fontsize in environment {verbatim}
\makeatletter
\def\verbatim{\small\@verbatim \frenchspacing\@vobeyspaces \@xverbatim}
%\def\verbatim@font{\small\ttfamily}
\makeatother

@@

\author{
  \vskip 1em
  Jiaxiang~Liu,
  Min~Zhou,
  Xiaoyu~Song, \emph{Senior Member}, \emph{IEEE},
  Ming~Gu,
  and Jiaguang~Sun

@@

    J.\ Liu is with the Department of Computer Science \& Technology,
    Tsinghua University, Beijing, China 
    \hide{, and also with LIX,
    \'Ecole Polytechnique, Palaiseau 91120, France} (e-mail:
    jiaxiang.liu@hotmail.com).

@@

Liu and Layland~\cite{DBLP:journals/jacm/LiuL73} gave a sufficient
condition for the schedulability of a set of $n$ tasks scheduled by
RMS: $\displaystyle\Sigma^n_{i=1}C_i/T_i \le n(2^{1/n}-1)$, where
$C_i$ and $T_i$ are the \emph{computation (time) requirement} and the
period of task $\tau_i$, respectively. Two main directions on RMS have
been explored since then. One is to relax the assumptions on the
original RMS model, making it applicable on more systems.  For
instance,
\cite{DBLP:conf/rtss/LehoczkySS87,DBLP:journals/rts/SpruntSL89,DBLP:conf/rtss/LehoczkyR92,DBLP:journals/tc/StrosniderLS95}
allow aperiodic tasks in the scheduling,
\cite{DBLP:journals/pe/LeungW82,audsley1993deadline} generalize RMS to
be \emph{deadline-monotonic}, \cite{DBLP:journals/tc/ShaRL90} allows
resource sharing among tasks,
\cite{dhall1978real,DBLP:journals/rts/LopezGDG03,DBLP:journals/tpds/LopezDG04,DBLP:journals/tc/BaruahG03}
extend RMS on multiprocessors, and
\cite{DBLP:journals/rts/OhS94,DBLP:journals/rts/GhoshMMS98,DBLP:journals/tpds/BertossiMR99}
enhance fault-tolerance. The other direction is to generate better
schedulability test conditions for the algorithm and its
extensions~\cite{DBLP:conf/rtss/LehoczkySD89,DBLP:conf/rtss/KuoM91,DBLP:journals/tc/BiniBB03,DBLP:journals/rts/LopezGDG03,DBLP:journals/tc/BaruahG03,gardner1999}. The
RMS algorithm is no doubt of practical importance.

@@

In this paper, we use \emph{Real-Time Maude}, a \emph{rewriting}-based
formal modeling language and analysis tool for real-time systems, to
model a realistic implementation of RMS that serves as a simplified
operating system within an avionic control system, and then verify
several desired properties. Based on a realistic implementation, our
model extends the standard RMS model proposed
in~\cite{DBLP:journals/jacm/LiuL73}, by considering overhead and other
details of the hardware platform.

@@

\subsection{Rate-Monotonic Scheduling Algorithm}
\label{ss:rms}
A task set consists of \emph{only} $n$ periodic tasks
$\tau_1,\ldots,\tau_n$. Each task $\tau_i$ has a period $T_i$ and a
computation requirement $C_i$. First jobs of all tasks are assumed to
be initiated at time $0$ simultaneously.  Deadlines consist of
runnability constraints only: the deadline of a job corresponding to
$\tau_i$ is the initiation time of the next job corresponding to
$\tau_i$.  The RMS algorithm chooses the labeling such that $T_1\le
T_2\le \ldots \le T_n$. Consequently, $\tau_i$ receives priority $i$,
assuming smaller numbers have higher priorities. The following
assumptions are made:

@@

A simple example showing this algorithm is depicted in
Fig.~\ref{f:example}(a). However in this paper, we consider an
implementation instead of the RMS algorithm itself, thus the model
would be more complicated than this standard, ideal setting.
%% Assumption
(A1) will be modified because of the interrupt mask
mechanism, while (A4) is relaxed to obtain a more realistic analysis
model.

@@

In this paper, we only consider Real-Time Maude's \emph{Linear
  Temporal Logic (LTL) model checker}, which analyzes whether
\emph{each} behavior satisfies a temporal logic formula. \emph{State
  propositions} are defined as terms of sort \verb|Prop|. Their
semantics is defined by conditional equations of the form $\texttt{ceq
} \mathit{statePattern} \texttt{ |= } \mathit{prop} \texttt{ = } b \texttt{ if } cond$,
with $b$ a term of sort \verb|Bool|, stating that $\mathit{prop}$ evaluates to
$b$ in states which are instances of $\mathit{statePattern}$ provided the
condition $cond$ holds. These equations together define $\mathit{prop}$ to hold
in all states $s$ that make $s \texttt{ |= } \mathit{prop}$ evaluate to
\verb|true|. A temporal logic \emph{formula} is constructed by state
propositions and temporal logic operators such as \verb|~|(negation),
\verb|\/|(disjunction), \verb|[]|(``always''),
\verb|U|(``until''). Real-Time Maude supports both \emph{timed} and
\emph{untimed LTL model checking}. The untimed model checking command
%% \indent
%% $\texttt{(mc } s \texttt{ |=u } \Phi \texttt{ .)}$\\
\begin{alltt}
  (mc \(s\) |=u \(\Phi\) .)
\end{alltt}
checks whether the temporal logic formula $\Phi$ holds in all
behaviors starting from the initial state $s$, \emph{with no time
  limit}.

@@

\begin{figure}[!t]
\centering
\includegraphics[width=8.8cm]{rms-pseudocode.pdf}
\caption{The C-Like Pseudo-code of $schedule()$}
\label{f:schedule}
\end{figure}

@@

\begin{figure}[!t]
\centering
\includegraphics{rms.pdf}
\caption{A set of $2$ periodic tasks scheduled under RMS algorithm and
  the implementation, respectively. $T=T_1=10$, $T_2=20$, $C_1=3$,
  $C_2=2$. In~(b), we assume both scheduling time and switching time
  are $2$.}
\label{f:example}
\end{figure}

@@

We also need
%a sort \verb|Interval| to denote computation requirements of tasks, and 
a sort \verb|Counter| to record the execution of tasks.  We call
\emph{execution time} the time how long a task has been executing
for. A \verb|Counter| records the execution time and the computation
requirement of a task.

@@

On the other hand, a system task is an object instance of the subclass
\verb|SysTask| of \verb|Task| with no extra attributes.
\hide{
\begin{verbatim}
  class SysTask | .  
  subclass SysTask < Task .
\end{verbatim}
}
Different from periodic tasks, system tasks are organized in a
multiset of sort \verb|SysTasks|, and identified by their \verb|Oid|s.

@@

When the execution time of the system task \verb|scheduling| reaches
its computation requirement, scheduling is finished. We model this
instantaneous action with the following rule:
\begin{verbatim}
  crl [scheduling-finish] :
    (L T STS HW ISRC) 
    => (SYSTEM).finishScheduling
    if SYSTEM := (L T STS HW ISRC) 
       /\ (SYSTEM).running == scheduling 
       /\ scheduling isComplete?in STS .
\end{verbatim}
where function \verb|_.running| returns the current \verb|pc| value of
the system, and \verb|_isComplete?in_| checks whether the execution
time of the task reaches its computation requirement.

@@

We have analyzed our model in ten different scenarios, including both
realistic ones provided by our industrial partner and experimental
ones designed by ourselves, four of them are described below:
\begin{itemize}
\item Scenario~(i) with 2 tasks $\tau_1$ and $\tau_2$: $T_1=5ms$,
  $C_1=3ms$, $T_2=25ms$, $C_2=7ms$.
\item Scenario~(ii) with 2 tasks $\tau_1$ and $\tau_2$: $T_1=5ms$,
  $C_1=2ms$, $T_2=25ms$, $C_2=2.3ms$.
\item Scenario~(iii) with 3 tasks $\tau_1$, $\tau_2$, $\tau_3$:
  $T_1=5ms$, $C_1=2.7ms$, $T_2=10ms$, $C_2=2ms$, $T_3=25ms$,
  $C_3=3ms$.
\item Scenario~(iv) with 3 tasks $\tau_1$, $\tau_2$, $\tau_3$:
  $T_1\!=\!5ms$, $C_1\!=\!2.5ms$, $T_2\!=\!10ms$, $C_2\!=\!1.5ms$,
  $T_3\!=\!15ms$, $C_3\!=\!4.5ms$.
\end{itemize}
Note that thanks to the expressiveness of Real-Time Maude, we
only need to define an initial state of sort \verb|System| to specify
a given task set. No necessity to modify the model is needed.

\begin{figure*}[!t]
\centering
\includegraphics{ce.pdf}
\caption{A counterexample of schedulability in Scenario~(iv). The first
  job instance of $\tau_3$ misses its deadline at time $15ms$, which
  is the initiation time for the second job instance of $\tau_3$.}
\label{f:counterexample}
\end{figure*}

Instantiating our model on dense time domain and choosing the
\emph{maximal time sampling strategy}, the results of the model
checking show that correctness property holds in all scenarios. 
%% As to the 
Schedulability property holds in Scenarios~(i--iii), but
fails in Scenario~(iv).  One counterexample of the schedulability
within Scenario~(iv), returned by the model checking command, is
pictured in Fig.~\ref{f:counterexample}, where
$\tau_3$ misses its deadline at time $15ms$.

@@

In general, completeness of untimed model checking cannot be achieved
for any systems, any time sampling strategies and any
properties. However, \"Olveczky and Meseguer proved the completeness
of untimed temporal logic model checking, under the maximal time
sampling strategy, for a large class of real-time systems possessing a
set of ``good'' properties that is called \emph{time-robustness}, and
for a set of ``good'' LTL formulae constructed by
\emph{tick-invariant} propositions\footnote{We avoid introducing the
  definitions of time-robustness and tick-invariance, due to the
  requirements of extra rewriting logic
  background.}~\cite{DBLP:journals/entcs/OlveczkyM07a}:
\begin{theorem}[\cite{DBLP:journals/entcs/OlveczkyM07a}]
\label{t:completeness}
Given a time-robust real-time rewrite theory $\cal R$, a set $AP$ of
tick-invariant atomic propositions, an LTL formula $\Phi$ (excluding
the \emph{next} operator $\bigcirc$) whose atomic propositions are
contained in $AP$. The untimed temporal logic model checking verifying
$\Phi$ is \emph{complete} under the maximal time sampling strategy.
\end{theorem}

@@

Considering schedulability test, Liu and
Layland~\cite{DBLP:journals/jacm/LiuL73} gave the famous sufficient
condition that a set of periodic tasks is schedulable with respect to
RMS if $\displaystyle \Sigma^n_{i=1} C_i/T_i \le n(2^{1/n}-1)$
holds. Then a more sufficient condition, known as \emph{Hyperbolic
  Bound}, which has the same complexity as Liu and Layland's, was
proposed in~\cite{DBLP:journals/tc/BiniBB03}. On the other hand,
necessary and sufficient conditions for schedulability were derived
independently in~\cite{DBLP:journals/rts/SpruntSL89}
and~\cite{audsley1993deadline}, requiring more sophisticated analysis
on the task set. Nevertheless, all these results take no overhead into
account, being not as realistic as ours. Katcher et al.\ did consider
overhead in their schedulability analysis, under several models based
on different kinds of popular
implementations~\cite{DBLP:journals/tse/KatcherAS93}.  However, our
target implementation is not in their scope. Furthermore, compared
with those theoretical analyses, our approach based on formal modeling
and verification has three advantages. One is that if our
schedulability test answers ``no'', it returns at the same time a real
counterexample, which is able to guide our engineer to adjust the
design, by changing either the priorities or even the scheduling
algorithm. The second is that, when a fresh scheduling strategy is
applied, our analysis can be adjusted only by modifying the model,
while theoretical approaches may need thorough analyses and reasoning.
The last one is that, considering overhead and details of hardware does
introduce some kind of non-determinism into the model. For example, in
our model, if the running task is complete \emph{right} at the time
when an interrupt request occurs, two different behaviors are
possible: (i)~the system performs task switching, during which the
interrupt request is masked, hence the scheduling and initiation of
tasks may be delayed; (ii)~the system answers the interrupt request
immediately, the switching being delayed. Tackling such
non-determinism via theoretical analyses seems more complicated than
our automatic approach.

@@

Time-robustness is a set of properties expected from a well-behaved
real-time rewrite theory. We avoid introducing the accurate definition
of time-robustness, but instead give the following lemma to prove
time-robustness:
\begin{lemma}[\cite{DBLP:journals/entcs/OlveczkyM07a}]
\label{l:timerobustness}
Let $\cal R$ be an object-oriented specification with a standard tick
rule, and let the infinity element \verb|INF| be the only element in
the time domain which is not a normal time value.
%and let the time domain be \emph{linear}.
Then $\cal R$ is time-robust if the following conditions are satisfied
for all appropriate ground terms $t$ and $r,r'$:
\begin{enumerate}
\item
[(i)] \verb|mte(delta(|$t$\verb|,|$r$\verb|))| $=$
\verb|mte(|$t$\verb|) monus |$r$, for all $r\le$
\verb|mte(|$t$\verb|)|, where \verb|monus| is the built-in minus
operation defined on sort \verb|Time|;

@@

Several auxiliary lemmas are proved:
\begin{lemma}
\label{l:auxtask}
Given $\mathit{ID}$ of sort \verb|MaybeNat| and $L$ of sort
\verb|TaskList|, $\mteTask{\mathit{ID}}{\deltaTask{\mathit{ID}}{L}{r}}
= \mteTask{\mathit{ID}}{L}\; \verb|monus|\;r$ for all
$r\le\mteTask{\mathit{ID}}{L}$.
\end{lemma}
\begin{IEEEproof}
If $\mathit{ID}=\verb|none|$, the case is trivial. By definition,
$\mteTask{\texttt{none}}{\deltaTask{\texttt{none}}{L}{r}} =
\mteTask{\texttt{none}}{L} = \verb|INF| =
\mteTask{\texttt{none}}{L}\\ \verb|monus|\; r$. Otherwise,
$\mathit{ID}= \verb|some|~N$ with $N$ of sort \verb|Nat|. Assume that
the \verb|cnt| of the $N$th task in $L$ is
\verb|[|$r_e$\verb|/|$C$\verb|]|. Then by definition,
$\deltaTask{\mathit{ID}}{L}{r}=L'$, where the
$N$th task in $L'$ has \verb|cnt| value being
\verb|[|$(r_e+r)$\verb|/|$C$\verb|]|. Hence,
\begin{IEEEeqnarray*}{Cl}
  & \mteTask{ID}{\deltaTask{ID}{L}{r}}
\\  
= & \mteTask{ID}{L'}
\\
= & C~\verb|monus|~(r_e+r) 
= (C~\verb|monus|~r_e)~\verb|monus|~r 
\\
= & \mteTask{ID}{L}~\verb|monus|~r\;. \hfill\IEEEQEDhere
\end{IEEEeqnarray*}
\end{IEEEproof}
\begin{lemma}
\label{l:auxis}
Given $\mathit{ISRC}$ of class \verb|IntSrc| representing an
reasonable state of interrupt source in our model,
$\mteIS{\deltaIS{\mathit{ISRC}}{r}} = \mteIS{\mathit{ISRC}}\;
\verb|monus|\\ r$ for all $r\le\mteIS{\mathit{ISRC}}$.
\end{lemma}
\begin{IEEEproof}
A reasonable $\mathit{ISRC}$ must be of the form
\IntSrc{$O$}{$v$}{$T$} with $v\le T$. Thus,
\begin{IEEEeqnarray*}{Cl}
  & \mteIS{\deltaIS{\mathit{ISRC}}{r}}
\\  
= & \mteIS{\IntSrc{O}{(v~\texttt{monus}~r)}{T}}
\\
= & v~\texttt{monus}~r
= \mteIS{\mathit{ISRC}}~\texttt{monus}~r\;. \hfill\IEEEQEDhere
\end{IEEEeqnarray*}
\end{IEEEproof}
\begin{lemma}
\label{l:auxhw}
Given $\mathit{HW}$ of sort \verb|Hardware|, for all
$r\le\mteIr{\mathit{HW}}$, $\mteIr{\mathit{HW}} =
\mteIr{\mathit{HW}}\; \verb|monus|\; r$.
\end{lemma}
\begin{IEEEproof}
  It concludes by discussing on whether there exists an interrupt
  request detected in $\mathit{HW}$.
\end{IEEEproof}

Now we can present the detailed proof of Theorem~\ref{t:main}:
\noindent
\begin{IEEEproof}[Proof of Theorem~\ref{t:main}]
We first prove the time-robustness of our model by
Lemma~\ref{l:timerobustness}.  Instantiated on the built-in dense time
domain \verb|POSRAT-TIME-DOMAIN-WITH-INF|, our model has \verb|INF| as
the only element which is not a normal time value. As presented in
Section~\ref{ss:timedbehavior}, only a single standard tick rule is
defined. Hence our model is time-robust by
Lemma~\ref{l:timerobustness} provided conditions (i--iv) hold.

@@

%% \section*{Acknowledgment}
%% The authors would like to thank the reviewers and editors for their
%% invaluable comments and suggestions.

@@

\bibliographystyle{Bibliography/IEEEtranTIE}
\bibliography{Bibliography/IEEEabrv,Bibliography/main}\ 
%IEEEabrv instead of IEEEfull

>>>>

%\documentclass[onecolumn]{IEEEtranTIE}
\documentclass[journal]{IEEEtranTIE}
\usepackage{graphicx}
\usepackage[noadjust]{cite}
\usepackage{picinpar}
\usepackage{amsmath}
\usepackage{stfloats}
\usepackage{url}
\usepackage{flushend}
\usepackage[latin1]{inputenc}
\usepackage{colortbl}
\usepackage{soul}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{color}
\usepackage{alltt}
\usepackage[hidelinks]{hyperref}
\usepackage{enumerate}
\usepackage{siunitx}
\usepackage{breakurl}
\usepackage{epstopdf}
\usepackage{pbox}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}

% Define the fontsize in environment {verbatim}
\makeatletter
\def\verbatim{\small\@verbatim \frenchspacing\@vobeyspaces \@xverbatim}
\makeatother

@@

\author{
  \vskip 1em
  Jiaxiang~Liu,
  Min~Zhou,
  Xiaoyu~Song,
  Ming~Gu,
  and Jiaguang~Sun

@@

    J.\ Liu is with the Department of Computer Science \& Technology,
    Tsinghua University, Beijing, China (e-mail:
    jiaxiang.liu@hotmail.com).

@@

Liu and Layland~\cite{DBLP:journals/jacm/LiuL73} gave a sufficient
condition for the schedulability of a set of $n$ tasks scheduled by
RMS: $\Sigma^n_{i=1}C_i/T_i \le n(2^{1/n}-1)$, where $C_i$ and $T_i$
are the \emph{computation (time) requirement} and the period of task
$\tau_i$, respectively. Two main directions on RMS have been explored
since then. One is to relax the assumptions on the original RMS model,
making it applicable on more systems.  For instance,
\cite{DBLP:conf/rtss/LehoczkySS87,DBLP:journals/rts/SpruntSL89,DBLP:conf/rtss/LehoczkyR92,DBLP:journals/tc/StrosniderLS95}
allow aperiodic tasks in the scheduling,
\cite{DBLP:journals/pe/LeungW82,audsley1993deadline} generalize RMS to
be \emph{deadline-monotonic}, \cite{DBLP:journals/tc/ShaRL90} allows
resource sharing among tasks,
\cite{dhall1978real,DBLP:journals/rts/LopezGDG03,DBLP:journals/tpds/LopezDG04,DBLP:journals/tc/BaruahG03}
extend RMS on multiprocessors, and
\cite{DBLP:journals/rts/OhS94,DBLP:journals/rts/GhoshMMS98,DBLP:journals/tpds/BertossiMR99}
enhance fault-tolerance. The other direction is to generate better
schedulability test conditions for the algorithm and its
extensions~\cite{DBLP:conf/rtss/LehoczkySD89,DBLP:conf/rtss/KuoM91,DBLP:journals/tc/BiniBB03,DBLP:journals/rts/LopezGDG03,DBLP:journals/tc/BaruahG03,gardner1999}. The
RMS algorithm is no doubt of practical importance.

@@

In this paper, we use \emph{Real-Time Maude}, a \emph{rewriting}-based
formal modeling language and analysis tool for real-time systems, to
model a realistic implementation of RMS that serves as a simplified
operating system within an avionic control system, and then verify
several desired properties. Based on a realistic implementation, our
model extends the standard RMS model proposed
in~\cite{DBLP:journals/jacm/LiuL73}, by taking into account overhead
and other details of the hardware platform.

@@

\subsection{Rate-Monotonic Scheduling Algorithm}
\label{ss:rms}
A task set consists of \emph{only} $n$ periodic tasks
$\tau_1,\ldots,\tau_n$. Each task $\tau_i$ has a period $T_i$ and a
computation requirement $C_i$. First jobs of all tasks are assumed to
be initiated at time $0$ simultaneously.  Deadlines consist of
runnability constraints only: the deadline of a job corresponding to
$\tau_i$ is the initiation time of the next job corresponding to
$\tau_i$.  The RMS algorithm chooses the labeling such that $T_1\le
T_2\le \ldots \le T_n$. Consequently, $\tau_i$ is given priority $i$,
assuming smaller numbers have higher priorities. The following
assumptions are made:

@@

A simple example showing this algorithm is depicted in
Fig.~\ref{f:example}(a). However in this paper, we consider an
implementation instead of the RMS algorithm itself, thus the model
would be more complicated than this standard, ideal setting.  (A1)
will be modified because of the interrupt mask mechanism, while (A4)
is relaxed to obtain a more realistic analysis model.

@@

In this paper, we only consider Real-Time Maude's \emph{Linear
  Temporal Logic (LTL) model checker}, which analyzes whether
\emph{each} behavior satisfies a temporal logic formula. \emph{State
  propositions} are defined as terms of sort \verb|Prop|. Their
semantics is defined by conditional equations of the form $\texttt{ceq
} \mathit{statePattern} \texttt{ |= } \mathit{prop} \texttt{ = } b
\texttt{ if } cond$, with $b$ a term of sort \verb|Bool|, stating that
$\mathit{prop}$ evaluates to $b$ in states which are instances of
$\mathit{statePattern}$ provided the condition $cond$ holds. These
equations together define $\mathit{prop}$ to hold in all states $s$
that make $s \texttt{ |= } \mathit{prop}$ evaluate to \verb|true|. A
temporal logic \emph{formula} is constructed by state propositions and
temporal logic operators such as \verb|~|(negation),
\verb|\/|(disjunction), \verb|[]|(``always''),
\verb|U|(``until''). Real-Time Maude supports both \emph{timed} and
\emph{untimed LTL model checking}. The untimed model checking command
\begin{alltt}
  (mc \(s\) |=u \(\Phi\) .)
\end{alltt}
checks whether the temporal logic formula $\Phi$ holds in all
behaviors starting from the initial state $s$, \emph{with no time
  limit}.

@@

\begin{figure}[!t]
\centering
\includegraphics[width=8.8cm]{FIG1_15-TIE-3480.pdf}
\caption{The C-Like Pseudocode of $schedule()$}
\label{f:schedule}
\end{figure}

@@

\begin{figure}[!t]
\centering
\includegraphics{FIG2_15-TIE-3480.pdf}
\caption{A set of $2$ periodic tasks scheduled under RMS algorithm and
  the implementation, respectively. $T=T_1=10$, $T_2=20$, $C_1=3$,
  $C_2=2$. In~(b), we assume both scheduling time and switching time
  are $2$.}
\label{f:example}
\end{figure}

@@

We also need a sort \verb|Counter| to record the execution of tasks.
We call \emph{execution time} the time how long a task has been
executing for. A \verb|Counter| records the execution time and the
computation requirement of a task.

@@

On the other hand, a system task is an object instance of the subclass
\verb|SysTask| of \verb|Task| with no extra attributes.  Different
from periodic tasks, system tasks are organized in a multiset of sort
\verb|SysTasks|, and identified by their \verb|Oid|s.

@@

When the execution time of the system task \verb|scheduling| reaches
its computation requirement, scheduling is finished. We model this
instantaneous action with the following rule:
\begin{verbatim}
  crl [scheduling-finish] :
    SYSTEM => (SYSTEM).finishScheduling
    if SYSTEM := (L T STS HW ISRC) 
       /\ (SYSTEM).running == scheduling 
       /\ scheduling isComplete?in STS .
\end{verbatim}
where function \verb|_.running| returns the current \verb|pc| value of
the system, and \verb|_isComplete?in_| checks whether the execution
time of the task reaches its computation requirement.

@@

We have analyzed our model in ten different scenarios, including both
realistic ones provided by our industrial partner and experimental
ones designed by ourselves, four of them are described below:
\begin{itemize}
\item Scenario~(i) with 2 tasks $\tau_1$ and $\tau_2$: $T_1=5ms$,
  $C_1=3ms$, $T_2=25ms$, $C_2=7ms$.
\item Scenario~(ii) with 2 tasks $\tau_1$ and $\tau_2$: $T_1=5ms$,
  $C_1=2ms$, $T_2=25ms$, $C_2=2.3ms$.
\item Scenario~(iii) with 3 tasks $\tau_1$, $\tau_2$, $\tau_3$:
  $T_1=5ms$, $C_1=2.7ms$, $T_2=10ms$, $C_2=2ms$, $T_3=25ms$,
  $C_3=3ms$.
\item Scenario~(iv) with 3 tasks $\tau_1$, $\tau_2$, $\tau_3$:
  $T_1=5ms$, $C_1=2.5ms$, $T_2=10ms$, $C_2=1.5ms$,
  $T_3=15ms$, $C_3=4.5ms$.
\end{itemize}
Note that thanks to the expressiveness of Real-Time Maude, we
only need to define an initial state of sort \verb|System| to specify
a given task set. No necessity to modify the model is needed.

\begin{figure*}[!t]
\centering
\includegraphics{FIG3_15-TIE-3480.pdf}
\caption{A counterexample of schedulability in Scenario~(iv). The first
  job instance of $\tau_3$ misses its deadline at time $15ms$, which
  is the initiation time for the second job instance of $\tau_3$.}
\label{f:counterexample}
\end{figure*}

Instantiating our model on dense time domain and choosing the
\emph{maximal time sampling strategy}, the results of the model
checking show that correctness property holds in all scenarios.
Schedulability property holds in Scenarios~(i--iii), but fails in
Scenario~(iv).  One counterexample of the schedulability within
Scenario~(iv), returned by the model checking command, is pictured in
Fig.~\ref{f:counterexample}, where $\tau_3$ misses its deadline at
time $15ms$.

@@

In general, completeness of untimed model checking cannot be achieved
for any systems, any time sampling strategies and any
properties. However, \"Olveczky and Meseguer proved the completeness
of untimed temporal logic model checking, under the maximal time
sampling strategy, for a large class of real-time systems possessing a
set of ``good'' properties that is called \emph{time-robustness}, and
for a set of ``good'' LTL formulae constructed by
\emph{tick-invariant} propositions\footnote{We avoid introducing the
  definitions of time-robustness and tick-invariance, due to the
  requirements of extra rewriting logic
  background.}~\cite{DBLP:journals/entcs/OlveczkyM07a}:
\begin{theorem}[\cite{DBLP:journals/entcs/OlveczkyM07a}]
\label{t:completeness}
Given a time-robust real-time rewrite theory $\mathcal{R}$, a set $AP$
of tick-invariant atomic propositions, an LTL formula $\Phi$
(excluding the \emph{next} operator $\bigcirc$) whose atomic
propositions are contained in $AP$. The untimed temporal logic model
checking verifying $\Phi$ is \emph{complete} under the maximal time
sampling strategy.
\end{theorem}

@@

Considering schedulability test, Liu and
Layland~\cite{DBLP:journals/jacm/LiuL73} gave the famous sufficient
condition that a set of periodic tasks is schedulable with respect to
RMS if $\Sigma^n_{i=1} C_i/T_i \le n(2^{1/n}-1)$ holds. Then a more
sufficient condition, known as \emph{Hyperbolic Bound}, which has the
same complexity as Liu and Layland's, was proposed
in~\cite{DBLP:journals/tc/BiniBB03}. On the other hand, necessary and
sufficient conditions for schedulability were derived independently
in~\cite{DBLP:journals/rts/SpruntSL89} and~\cite{audsley1993deadline},
requiring more sophisticated analysis on the task set. Nevertheless,
all these results take no overhead into account, being not as
realistic as ours. Katcher et al.\ did consider overhead in their
schedulability analysis, under several models based on different kinds
of popular implementations~\cite{DBLP:journals/tse/KatcherAS93}.
However, our target implementation is not in their scope. Furthermore,
compared with those theoretical analyses, our approach based on formal
modeling and verification has three advantages. One is that if our
schedulability test answers ``no'', it returns at the same time a real
counterexample, which is able to guide our engineer to adjust the
design, by changing either the priorities or even the scheduling
algorithm. The second is that, when a fresh scheduling strategy is
applied, our analysis can be adjusted only by modifying the model,
while theoretical approaches may need thorough analyses and reasoning.
The last one is that, considering overhead and details of hardware
does introduce some kind of non-determinism into the model. For
example, in our model, if the running task is complete \emph{right} at
the time when an interrupt request occurs, two different behaviors are
possible: (i)~the system performs task switching, during which the
interrupt request is masked, hence the scheduling and initiation of
tasks may be delayed; (ii)~the system answers the interrupt request
immediately, the switching being delayed. Tackling such
non-determinism via theoretical analyses seems more complicated than
our automatic approach.

@@

Time-robustness is a set of properties expected from a well-behaved
real-time rewrite theory. We avoid introducing the accurate definition
of time-robustness, but instead give the following lemma to prove
time-robustness:
\begin{lemma}[\cite{DBLP:journals/entcs/OlveczkyM07a}]
\label{l:timerobustness}
Let $\mathcal{R}$ be an object-oriented specification with a standard
tick rule, and let the infinity element \verb|INF| be the only element
in the time domain which is not a normal time value.  Then
$\mathcal{R}$ is time-robust if the following conditions are satisfied
for all appropriate ground terms $t$ and $r,r'$:
\begin{enumerate}
\item
[(i)] \verb|mte(delta(|$t$\verb|,|$r$\verb|))| $=$
\verb|mte(|$t$\verb|) monus |$r$, for all $r\le$
\verb|mte(|$t$\verb|)|, where \verb|monus| is the built-in minus
operation defined on sort \verb|Time|;

@@

Several auxiliary lemmas are proved:
\begin{lemma}
\label{l:auxtask}
Given $\mathit{ID}$ of sort \verb|MaybeNat| and $L$ of sort
\verb|TaskList|, $\mteTask{\mathit{ID}}{\deltaTask{\mathit{ID}}{L}{r}}
= \mteTask{\mathit{ID}}{L}\; \verb|monus|\;r$ for all
$r\le\mteTask{\mathit{ID}}{L}$.
\end{lemma}
\begin{IEEEproof}
If $\mathit{ID}=\verb|none|$, the case is trivial. By definition,
$\mteTask{\texttt{none}}{\deltaTask{\texttt{none}}{L}{r}} =
\mteTask{\texttt{none}}{L} = \verb|INF| =
\mteTask{\texttt{none}}{L}\\ \verb|monus|\; r$. Otherwise,
$\mathit{ID}= \verb|some|~N$ with $N$ of sort \verb|Nat|. Assume that
the \verb|cnt| of the $N$th task in $L$ is
\verb|[|$r_e$\verb|/|$C$\verb|]|. Then by definition,
$\deltaTask{\mathit{ID}}{L}{r}=L'$, where the
$N$th task in $L'$ has \verb|cnt| value being
\verb|[|$(r_e+r)$\verb|/|$C$\verb|]|. Hence,
\begin{IEEEeqnarray*}{Cl}
  & \mteTask{\mathit{ID}}{\deltaTask{\mathit{ID}}{L}{r}}
\\  
= & \mteTask{\mathit{ID}}{L'}
\\
= & C~\verb|monus|~(r_e+r) 
= (C~\verb|monus|~r_e)~\verb|monus|~r 
\\
= & \mteTask{\mathit{ID}}{L}~\verb|monus|~r\;. \hfill\IEEEQEDhere
\end{IEEEeqnarray*}
\end{IEEEproof}

\begin{lemma}
\label{l:auxis}
Given $\mathit{ISRC}$ of class \verb|IntSrc| representing an
reasonable state of interrupt source in our model,
$\mteIS{\deltaIS{\mathit{ISRC}}{r}} = \mteIS{\mathit{ISRC}}\;
\verb|monus|\\ r$ for all $r\le\mteIS{\mathit{ISRC}}$.
\end{lemma}
\begin{IEEEproof}
A reasonable $\mathit{ISRC}$ must be of the form
$\IntSrc{O}{v}{T}$ with $v\le T$. Thus,
\begin{IEEEeqnarray*}{Cl}
  & \mteIS{\deltaIS{\mathit{ISRC}}{r}}
\\  
= & \mteIS{\IntSrc{O}{(v~\texttt{monus}~r)}{T}}
\\
= & v~\texttt{monus}~r
= \mteIS{\mathit{ISRC}}~\texttt{monus}~r\;. \hfill\IEEEQEDhere
\end{IEEEeqnarray*}
\end{IEEEproof}

\begin{lemma}
\label{l:auxhw}
Given $\mathit{HW}$ of sort \verb|Hardware|, for all
$r\le\mteIr{\mathit{HW}}$, $\mteIr{\mathit{HW}} =
\mteIr{\mathit{HW}}\; \verb|monus|\; r$.
\end{lemma}
\begin{IEEEproof}
  It concludes by discussing on whether there exists an interrupt
  request detected in $\mathit{HW}$.
\end{IEEEproof}

Now we can present the detailed proof of Theorem~\ref{t:main}:
\begin{IEEEproof}[Proof of Theorem~\ref{t:main}]
We first prove the time-robustness of our model by
Lemma~\ref{l:timerobustness}.  Instantiated on the built-in dense time
domain \verb|POSRAT-TIME-DOMAIN-WITH-INF|, our model has \verb|INF| as
the only element which is not a normal time value. As presented in
Section~\ref{ss:timedbehavior}, only a single standard tick rule is
defined. Hence our model is time-robust by
Lemma~\ref{l:timerobustness} provided conditions (i--iv) hold.

@@

@@

\bibliographystyle{Bibliography/IEEEtranTIE}
\bibliography{Bibliography/IEEEabrv,Bibliography/BIB_15-TIE-3480}\
